# Reproducible Research: Peer Assessment 1

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
```   

## Loading and Pre-processing the Data
### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
#### 2. Process/transform the data if necessary into a format suitable for your analysis
```{r,}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```
------

## What is mean total number of steps taken by day?
```{r,}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```
#### 1. Histogram of the total number of steps taken each day

```{r}
 qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

#### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMean <- (stepsByDay)
stepsByDayMedian <- (stepsByDay)
```
* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

-----

## What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot of the average number of steps taken
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
geom_line() +
xlab("5-minute interval") + 
ylab("average number of steps taken")
```
##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
* Most Steps at: `r timeMostSteps`

## Imputing the missing values
#### 1. Calculate and report the total number of missing values in the dataset
```{r}
numMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numMissingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
if(!require('imputeR')) 
  install.packages('imputeR')
  library('imputeR')
```

```{r}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

#### 4. Make a historgram of the total number of steps taken each day after missing values are imputed
```{r}
stepsByDayImputed <-tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

#### and Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayImputed`
* Median (Imputed): `r stepsByDayMedianImputed`

------

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
averages <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averages, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
  ylab("Number of steps")
```






  

          

 
 
 
 
 
 






